---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(dplyr)

intovalue_all <- readr::read_csv(here::here("data", "intovalue.csv"))
```

# IntoValue Dataset

## Overview

```{r filter-intovalue}
intovalue <-
  intovalue_all %>% 
  filter(
    has_german_umc_lead,
    !is_dupe,
    has_publication,
    has_pmid,
    has_pubmed,
    has_ft_pdf
  )
n_iv_trials_w_pubs <- nrow(intovalue)
```

We are interested in IntoValue trials with a German UMC lead and a publication with a PMID that resolves to a PubMed record and for which we could acquire the full-text as a PDF (n = `r n_iv_trials_w_pubs`).

```{r trials-same-pmid}
trials_same_pmid <- janitor::get_dupes(intovalue, pmid)

n_trials_same_pmid <- n_distinct(trials_same_pmid$id)
n_pmids_same_trial <- n_distinct(trials_same_pmid$pmid)
n_pmids_dupes <- unique(range(trials_same_pmid$dupe_count))
```

In general, there is max 1 publication per trial and max 1 trial per publication. However, there are `r n_trials_same_pmid` trials associated with the same `r n_pmids_same_trial` publications (i.e., `r n_pmids_dupes` publications per trial). Since the unit of analysis is trials, we disregard this double-counting of publications.

## TRN reporting in abstract

```{r trn-abs}
n_trn_abs <- nrow(filter(intovalue, has_trn_abstract))

prop_trn_abs <- n_trn_abs/n_iv_trials_w_pubs
```

<!-- $$ \text{TRN in abstract (%)} = \frac{\text{Number of trials with PubMed publications with TRN in abstract}}{\text{Number of trials with PubMed publications available as PDF full-text}}$$ -->

**Numerator**: Number of trials with PubMed publications with TRN in abstract

**Denominator**: Number of trials with PubMed publications available as PDF full-text

`r scales::percent(prop_trn_abs)` (`r n_trn_abs`/`r n_iv_trials_w_pubs`) of trials report a TRN in the abstract of their results publication.

## TRN reporting in full-text

```{r trn-ft}
n_trn_ft <- nrow(filter(intovalue, has_trn_ft_pdf))

prop_trn_ft <- n_trn_ft/n_iv_trials_w_pubs
```

**Numerator**: Number of trials with PubMed publications with TRN in PDF full-text

**Denominator**: Number of trials with PubMed publications available as PDF full-text

`r scales::percent(prop_trn_ft)` (`r n_trn_ft`/`r n_iv_trials_w_pubs`) of trials report a TRN in the full-text (PDF) of their results publication.

## Linked publication in registry

```{r reg-pub-link}
intovalue_ctgov <- filter(intovalue, is_c_tgov)

n_iv_trials_w_pubs_ctgov <- nrow(intovalue_ctgov)

n_reg_pub_link_ctgov <- nrow(filter(intovalue_ctgov, has_reg_pub_link))

prop_reg_pub_link_ctgov <- n_reg_pub_link_ctgov/ n_iv_trials_w_pubs_ctgov

filter(intovalue, is_c_tgov, 
       # has_reg_pub_link
       ) %>% filter(reference_type != "derived")

n_auto <- nrow(filter(intovalue_ctgov, reference_type == "derived"))
n_manual <- nrow(filter(intovalue_ctgov, reference_type != "derived"))
```

*Limitations*: We currently checked only clinicaltrials.gov for linked publications. DRKS is forthcoming if feasible. We consider a publication "linked" if the PMID is included in the trial registrations. Note that some publications are included in the registrations without a PMID.

**Numerator**: Number of trials with PubMed publications PMIDs linked in trial registration

**Denominator**: Number of trials (clinicaltrials.gov only) with PubMed publications available as PDF full-text

`r scales::percent(prop_reg_pub_link_ctgov)` (`r n_reg_pub_link_ctgov`/`r n_iv_trials_w_pubs_ctgov`) of trials on clinicaltrials.gov include a link (i.e., PMID) to their PubMed publication (as available in the IntoValue dataset). This include `r n_auto` (`r scales::percent(n_auto/n_reg_pub_link_ctgov)`) trials with automatically indexed publications (i.e., using TRN in PubMed's secondary identifier field) and `r n_manual` (`r scales::percent(n_manual/n_reg_pub_link_ctgov)`) trials with manually added publications.
